Computation and comparison of the stable northeastern United States marine boundary layer

### 2020-12-06 Paper/main.tex
Apply changes suggested by NREL Editorial office

\title{Computation and comparison of the stable Northeastern US marine boundary layer}

@@

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary
  layers located near the coast of the Northeastern United States.
  Using the ExaWind LES codes, three stable atmospheric conditions
  were chosen to match the Cape Wind measurements of Archer \emph{et
    al.} with wind speeds of 5 m/s, 10 m/s, and 15 m/s at the 20 m
  measurement height.  The behavior of the stable boundary layers,
  including mean flow quantities and turbulent statistics, are
  examined and compared to previous computations of the neutral and
  unstable offshore boundary layer at the same location.  This study
  also examines the domain and mesh requirements necessary to capture
  the turbulent scales for the stable offshore boundary layers.
  Finally, we compare solutions computed using both Nalu-Wind and
  AMR-Wind solvers, and report on their relative accuracy and
  performance in this study.

@@

\lettrine{T}he planned installation of several offshore wind energy
plants in the United States has highlighted the need to properly
understand the wind resource and atmospheric characteristics of the US
Atlantic Coast.  Several atmospheric phenomena particular to this
region, such as coastal low-level jets or seasonal Nor’easters, have
the potential to substantially impact the operation and power
production of offshore wind plants.  Of particular interest to the
current study is the atmospheric stability of the marine boundary
layer in the Northeastern US.

@@

Several recent measurement campaigns have provided data to understand
the ABL and wind characteristics for potential offshore wind farms in
the Northeastern US.  Pichugina et al. \cite{pichugina2017properties}
measured the wind profiles and vertical shear profiles in the Gulf of
Maine using a ship-borne lidar approach.  Analysis of measured data
sets from Nantucket Sound by Archer et
al. \cite{archer2016predominance} showed a predominance of low-shear,
unstable conditions at the site.  However, strong seasonal variations
and stratification changes due to diurnal variation were also
observed.

@@

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation
with a nonlinear SGS model containing backscatter to favorably compare
to the Beaufort Sea Arctic Stratus Experiment (BASE) measurements,
showing that the SBL could reach a quasi steady-state. They also
showed good agreement with the Nieustadt analytical model. In the same
year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at z=500 m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2m and 12.5m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25m resolution, but
very little at a 2m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieustadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

An inter-comparison of LES simulations of the SBL was organized in 2006, the
results of which are published in Beare et al.\ \cite{Beare2006}. Collectively,
this is known as the GABLS initiative, and involved simulating a moderately
stable flow taken from Kosovic and Curry \cite{Kosovic2000}. Eleven codes
were compared in all, wherein grid deltas of less than 12.5m were found to
successfully maintain resolved turbulence, grid deltas less than 3.125m
showed little statistical change, and results became independent of SGS
modeling at 1m resolution. In the same year, Kumar et al. \cite{Kumar2006}
successfully simulated an entire diurnal cycle with LES using a Lagrangian
dynamic scale dependent SGS model. This model has been used by several
researchers since 2011, including Lu and Porte-Agel \cite{Lu2011}, Huang
and Bou-Zeid \cite{Huang2013}, Park et al.\ \cite{Park2014}, and
Abkar and Porte-Agel \cite{Abkar2015}. Stoll and Porte-Agel \cite{Stoll2007}
compared three grid refinements and three scale dependent dynamic SGS models,
showing that the Lagrangian and locally averaged models reproduced important
flow features. The work of van Dop and Axelsen \cite{vanDop2007} provided
even more evidence in support Nieuwstadt's local scaling hypothesis for SBLs.
Chow and Zhou \cite{Chow2011} showed evidence that the Smagorinsky model is
simply too dissipative for use in strongly stable flows.

More recent work in LES of the SBL includes Matheou \cite{Matheou2016} who
explored a large parameter space using the Smagorinsky-Lilly SGS
closure, as well as Sullivan et al.\ \cite{sullivan2016turbulent}. Several researchers
have applied SBL simulation in the wind energy space \cite{Aitken2014, Abkar2015, Ghaisas2017}.
In 2020, Wurps et al. \cite{Wurps2020} provided a rigorous grid resolution study for the SBL,
looking at mean flow, turbulence variables, and two-point correlations, using a dynamic
SGS model that allows backscatter. They found grid requirements to be 2.5 m, 10 m, and 20 m
for stable, neutral, and convective boundary layers respectively. They advocate looking
at two point correlations in order to quantify flow structure size, which in turn
provides a basis for evaluating grid resolution. In all of the reviewed literature, SBL LES
has been run in the onshore context only, a complete comparison
including offshore stable ABL conditions has yet to be completed.

\subsection{Objectives}
In the current study on stable marine
boundary layers, we aim to build on the previous work in the
literature and complete several objectives.  First, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Secondly, two LES solvers
within the ExaWind project will be compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same SBL, and the relative differences between
the codes are reported.  Finally, a series of stable boundary layer
conditions will be computed using AMR-Wind which match the
measurements of Archer et al. \cite{archer2016predominance}.  The
behavior of the stable boundary layers, including mean flow quantities
and turbulent statistics, are examined and compared to the neutral and
unstable cases computed in \cite{cheung2020large}.

@@

\section{Conclusion}
In this study we have simulated the stable offshore boundary layer
matching the characteristics of measurements from Nantucket Sound in
the Northeastern US.  Three cases with wind speeds of 5 m/s, 10 m/s,
and 15 m/s were selected, and simulations were performed using the
ExaWind codes to match the TI and stability levels observed at from
the Cape Wind platform.  When examining the effect of stratification
using the Obukhov length, differences in the ABL behavior were seen
between the very stable boundary layer cases (5 m/s and 10 m/s), and
the stably stratified case (15 m/s).  The very stable cases had
shorter turbulent structures and less growth in the boundary layer
compared to the 15 m/s case.  The behavior of the three stable
boundary layers was also compared to the neutral and unstable cases
studied in \cite{cheung2020large}.  As expected, the turbulent
structures in all stable cases were noticeably smaller than in the
neutral and unstable boundary layers.

In addition to these simulations, we also examined the domain and mesh
requirements necessary for the stable offshore boundary layers
considered here.  Based on the calculated turbulent integral
lengthscale, horizontal domain sizes of 750 m $\times$ 750 m were found
to be necessary for these simulations.  Various mesh resolutions were
also considered for the stable runs, and it was determined that the
finest resolution of $\Delta_x$=2.5 m was able to resolve the longitudinal
length scales and highest temporal frequencies, but inadequate for
resolving the turbulence in the lateral direction.  Further study is
warranted to comprehensively determine the impact of stable
stratification on the required mesh resolution.

@@

This research was supported by the Wind Energy Technologies Office of
the US Department of Energy Office of Energy Efficiency and Renewable
Energy.  Sandia National Laboratories is a multimission laboratory
managed and operated by National Technology \& Engineering Solutions
of Sandia, LLC, a wholly owned subsidiary of Honeywell International
Inc., for the U.S. Department of Energy's National Nuclear Security
Administration under contract DE-NA0003525. The views expressed in the
article do not necessarily represent the views of the U.S. Department
of Energy or the United States Government.

@@

A portion of the research was performed using computational
resources sponsored by the Department of Energy's Office of Energy
Efficiency and Renewable Energy and located at the National Renewable
Energy Laboratory.

>>>>

\title{Computation and comparison of the stable northeastern United States
  marine boundary layer}

@@

\begin{abstract}
  In this work we investigate the behavior of stable marine boundary layers
  located near the coast of the Northeastern United States.  Using the ExaWind
  large eddy simulation (LES) codes, three stable atmospheric conditions were
  chosen to match the Cape Wind measurements of Archer \emph{et al.} with wind
  speeds of 5 m/s, 10 m/s, and 15 m/s at the 20 m measurement height.  The
  behavior of the stable boundary layers, including mean flow quantities and
  turbulent statistics, are examined and compared to previous computations of
  the neutral and unstable offshore boundary layer at the same location.  This
  study also examines the domain and mesh requirements necessary to capture the
  turbulent scales for the stable offshore boundary layers.  Finally, we compare
  solutions computed using both Nalu-Wind and AMR-Wind solvers, and report on
  their relative accuracy and performance in this study.

@@

\lettrine{T}he planned installation of several offshore wind energy
plants in the United States has highlighted the need to properly
understand the wind resource and atmospheric characteristics of the U.S.
Atlantic Coast.  Several atmospheric phenomena particular to this
region, such as coastal low-level jets or seasonal Nor’easters, have
the potential to substantially impact the operation and power
production of offshore wind plants.  Of particular interest to the
current study is the atmospheric stability of the marine boundary
layer in the Northeastern United States.

@@

Several recent measurement campaigns have provided data to understand
the ABL and wind characteristics for potential offshore wind farms in
the Northeastern United States.  Pichugina et al. \cite{pichugina2017properties}
measured the wind profiles and vertical shear profiles in the Gulf of
Maine using a ship-borne lidar approach.  Analysis of measured data
sets from Nantucket Sound by Archer et
al. \cite{archer2016predominance} showed a predominance of low-shear,
unstable conditions at the site.  However, strong seasonal variations
and stratification changes due to diurnal variation were also
observed.

@@

In 2000, Kosovic and Curry \cite{Kosovic2000} used an LES simulation
with a nonlinear SGS model containing backscatter to favorably compare
to the Beaufort Sea Arctic Stratus Experiment (BASE) measurements,
showing that the SBL could reach a quasi steady-state. They also
showed good agreement with the Nieuwstadt analytical model. In the same
year, Saiki et al. \cite{Saiki2000} simulated an SBL with
a strong capping inversion at z=500 m, producing reasonable surface layer profiles and
identifying a need for SBL specific SGS models. Beare and MacVean \cite{Beare2004}
presented the first grid-focused work on LES of the SBL, using grids with deltas
between 2 m and 12.5 m, finding significant variation between grid sizes. The inclusion
of a backscatter model resulted in significant variation at a 6.25 m resolution, but
very little at a 2 m resolution. Basu et al. \cite{Basu2005} published a set of
turbulence statistics, measurements, and LES simulations in support
of the viability of Nieuwstadt's local scaling hypothesis under very stable
conditions. Jimenez and Cuxart \cite{Jimenez2005} set out to quantify the range
for which standard SGS models, using the Kolmogorov theory for dissipation, were
valid for increasingly stable flows. They found the model to perform adequately
for weakly stable, and some moderately stable flow regimes.

An inter-comparison of LES simulations of the SBL was organized in 2006, the
results of which are published in Beare et al.\ \cite{Beare2006}. Collectively,
this is known as the GABLS initiative, and involved simulating a moderately
stable flow taken from Kosovic and Curry \cite{Kosovic2000}. Eleven codes
were compared in all, wherein grid deltas of less than 12.5 m were found to
successfully maintain resolved turbulence, grid deltas less than 3.125m
showed little statistical change, and results became independent of SGS
modeling at 1 m resolution. In the same year, Kumar et al. \cite{Kumar2006}
successfully simulated an entire diurnal cycle with LES using a Lagrangian
dynamic scale dependent SGS model. This model has been used by several
researchers since 2011, including Lu and Porte-Agel \cite{Lu2011}, Huang
and Bou-Zeid \cite{Huang2013}, Park et al.\ \cite{Park2014}, and
Abkar and Porte-Agel \cite{Abkar2015}. Stoll and Porte-Agel \cite{Stoll2007}
compared three grid refinements and three scale dependent dynamic SGS models,
showing that the Lagrangian and locally averaged models reproduced important
flow features. The work of van Dop and Axelsen \cite{vanDop2007} provided
even more evidence in support Nieuwstadt's local scaling hypothesis for SBLs.
Chow and Zhou \cite{Chow2011} showed evidence that the Smagorinsky model is
simply too dissipative for use in strongly stable flows.

More recent work in LES of the SBL includes Matheou \cite{Matheou2016} who
explored a large parameter space using the Smagorinsky-Lilly SGS
closure, as well as Sullivan et al.\ \cite{sullivan2016turbulent}. Several researchers
have applied SBL simulation in the wind energy space \cite{Aitken2014, Abkar2015, Ghaisas2017}.
In 2020, Wurps et al. \cite{Wurps2020} provided a rigorous grid resolution study for the SBL,
looking at mean flow, turbulence variables, and two-point correlations, using a dynamic
SGS model that allows backscatter. They found grid requirements to be 2.5 m, 10 m, and 20 m
for stable, neutral, and convective boundary layers respectively. They advocate looking
at two point correlations in order to quantify flow structure size, which in turn
provides a basis for evaluating grid resolution. In all of the reviewed literature, SBL LES
has been run in the onshore context only; a complete comparison
including offshore stable ABL conditions has yet to be completed.

\subsection{Objectives}
In the current study on stable marine
boundary layers, we aim to build on the previous work in the
literature and complete several objectives.  First, we examine the
domain and mesh requirements necessary to capture the turbulent scales
for the stable offshore boundary layers.  Secondly, two LES solvers
within the ExaWind project will be compared in this study: the
unstructured, incompressible Navier-Stokes solver Nalu-Wind and the
structured, adaptive mesh refinement solver AMR-Wind.  Both codes are
used to calculate the same SBL, and the relative differences between
the codes are reported.  Finally, a series of stable boundary layer
conditions will be computed using AMR-Wind that match the
measurements of Archer et al. \cite{archer2016predominance}.  The
behavior of the stable boundary layers, including mean flow quantities
and turbulent statistics, are examined and compared to the neutral and
unstable cases computed in \cite{cheung2020large}.

@@

\section{Conclusion}
In this study, we have simulated the stable offshore boundary layer matching the
characteristics of measurements from Nantucket Sound in the Northeastern United
States. Three cases with wind speeds of 5 m/s, 10 m/s, and 15 m/s were
selected, and simulations were performed using the ExaWind codes to match the TI
and stability levels observed at the Cape Wind platform.  When examining
the effect of stratification using the Obukhov length, differences in the ABL
behavior were seen between the very stable boundary layer cases (5 m/s and 10
m/s), and the stably stratified case (15 m/s).  The very stable cases had
shorter turbulent structures and less growth in the boundary layer compared to
the 15 m/s case.  The behavior of the three stable boundary layers was also
compared to the neutral and unstable cases studied in \cite{cheung2020large}.
As expected, the turbulent structures in all stable cases were noticeably
smaller than in the neutral and unstable boundary layers.

In addition to these simulations, we also examined the domain and mesh
requirements necessary for the stable offshore boundary layers
considered here.  Based on the calculated turbulent integral
length scale, horizontal domain sizes of 750 m $\times$ 750 m were found
to be necessary for these simulations.  Various mesh resolutions were
also considered for the stable runs, and it was determined that the
finest resolution of $\Delta_x$=2.5 m was able to resolve the longitudinal
length scales and highest temporal frequencies, but inadequate for
resolving the turbulence in the lateral direction.  Further study is
warranted to comprehensively determine the impact of stable
stratification on the required mesh resolution.

@@

This research was supported by the Wind Energy Technologies Office of the
U.S. Department of Energy (DOE) Office of Energy Efficiency and Renewable
Energy.  Sandia National Laboratories is a multimission laboratory managed and
operated by National Technology \& Engineering Solutions of Sandia, LLC, a
wholly owned subsidiary of Honeywell International Inc., for the U.S. Department
of Energy's National Nuclear Security Administration under contract
DE-NA0003525. The views expressed in the article do not necessarily represent
the views of the U.S. Department of Energy or the United States Government.

@@

A portion of the research was performed using computational resources sponsored
by the DOE's Office of Energy Efficiency and Renewable Energy and located at the
National Renewable Energy Laboratory.
